---
title: "FZahir_Assign15"
author: "Farhana Zahir"
date: "5/17/2020"
output:
  html_document:
    toc: true
    toc_float: yes
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r message=F, warning=F}
library(tidyverse)
```

## Problem 01

Find the equation of the regression line for the given points.
Round any final values to the nearest hundredth, if necessary. 

( 5.6, 8.8 ), ( 6.3, 12.4 ), ( 7, 14.8 ), ( 7.7, 18.2 ), ( 8.4, 20.8 )

Solution:

```{r}
df <- data.frame(x=c(5.6, 6.3, 7, 7.7, 8.4), y=c(8.8, 12.4, 14.8, 18.2, 20.8))
print (df)
```
Running the regression

```{r}
m1 = lm(y ~ x, data=df)
summary(m1)
```
Equation of the regression line:

$$y=−14.8+4.26∗x$$

Plotting the regression line

```{r}
ggplot(df, aes(x, y)) + 
  geom_point(colour="red", size=2) + 
  geom_abline(aes(slope=round(m1$coefficients[2], 2), intercept=round(m1$coefficients[1], 2))) +
  xlab("x values") + ylab("y values") + labs(title = "x vs. y")+
  theme(plot.title = element_text(hjust = 0.5))
```

## Problem 02

Find all local maxima, local minima, and saddle points for the function given below.
Write your answer(s) in the form (x, y, z). Separate multiple points with a comma.

$$f(x,y)=24x−6xy^2−8y^3$$

Solution:

$f_x(x,y)=\frac{∂}{∂x}(24x−6xy^2−8y^3)=24−6y^2$

$f_y(x,y)=\frac{∂}{∂y}(24x−6xy^2−8y^3)=−12xy−24y^2$

Setting $f_x(x,y)=0$,

$24−6y^2=0$ 

$6y^2=24$

y=±2 

Setting $f_y(x,y)=0$ and by substituting value of y,

$−12xy−24y^2=0$

$−12xy=24y^2$

x=−2y

When y=2, x=−4

When y=−2, x=4

Substituting these values in the function,

f(x,y)=f(−4,2)=(24x−6xy^2−8y^3)=[24(−4)−6(−4)(22)−8(2)3]=−64 

f(x,y)=f(4,−2)=(24x−6xy^2−8y^3)=[24(4)−6(4)(−22)−8(−2)3]=64

The critical points are (−4,2) and (4,−2)

(x,y,z) = (4,−2,64) and (−4,2,−64)

Second Derivative test:

Let $D=f_{xx}(x0,y0).f_{yy}(x0,y0)−f^2_{xy}(x0,y0)$
  
$f_{xx}(x,y)=\frac{∂}{∂x}(fx)=\frac{∂}{∂x}(24−6y^2)=0$

$f_{yy}(x,y)=\frac{∂}{∂y}(fy)=\frac{∂}{∂x}(−12xy−24y^2)=−12x−48y$

$f_{xy}(x,y)=\frac{∂}{∂y}(fx)=\frac{∂}{∂y}(24−6y^2)=−12y$

$D(x0,y0)=f_{xx}(x0,y0).f_{yy}(x0,y0)−f^2_{xy}(x0,y0)$

D(−4,2)=(0)((−12∗−4)−(48∗2))−(−12∗2)2=−576 

D(4,−2)=(0)((−12∗4)−(48∗(−2)))−(−12∗(−2))2=−576

D(−4,2)  and D(4,−2) are both negative, so (−4,2) and (4,−2) are the saddle points.


## Problem 03

A grocery store sells two brands of a product, the “house” brand and a “name” brand.
The manager estimates that if she
sells the “house” brand for x dollars and the “name” brand for y dollars, she will be able to sell 81−21x+17y units of the “house” brand and 40+11x−23y units of the “name” brand.

Step 1. Find the revenue function R(x, y).

Step 2.What is the revenue if she sells the “house” brand for $2.30 and the “name” brand for $4.10?

Solution:

If the manager sells the “house” brand for x dollars, she is able to sell 81−21x+17y
If the manager sells the “name” brand for y dollars, she is able to sell 40+11x−23y

$$Revenue(x)=x∗(81−21x+17y)=(81x−21x^2+17xy)$$

$$Revenue(y)=y∗(40+11x−23y)=(40y+11xy−23y^2)$$

$$R(x,y)=81x−21x^2+17xy+40y+11xy−23y^2 $$

$$R(x,y)=−21x^2+28xy−23y^2+81x+40y$$

Step 2:

x=$2.30 
y=$4.10

R(2.30,4.10)=−21(2.30)2+28(2.30∗4.10)−23(4.10)2+81(2.30)+40(4.10)

R(2.30,4.10)=$116.62

The total revenue if the manager sells the “house” brand for 2.30 and “name” brand for 4.10 is $116.62.


## Problem 04

A company has a plant in Los Angeles and a plant in Denver.
The firm is committed to produce a total of 96 units of a product each week. The total weekly cost is given by 

$$C(x,y)=16x^2+16y^2+7x+25y+700,$$ 
where x is the number of units produced in Los Angeles and y is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?


Solution:

Since the company is committed to produce a total of 96 units of a product each week, 

x+y=96

x=96−y

$C(96−y,y)=16(96−y)^2+16y^2+7(96−y)+25y+700 $

$C(96−y,y)=16(y^2−192y+9216)+16y^2+18y+1372$

$C(96−y,y)=13y^2−14y+2908$

To find the minimum value, we need to differentiate the function C(x,y) and equate it to 0.

$C′(96−y,y)=\frac{d}{dy}(13y^2−14y+2908)=23y−14$

23y−14=0

y=21

using value of y,

x= 96-21 = 75

The company needs to produce 75 units in Los Angeles plant and 21 units in Denver plant to minimize the total weekly cost.


## Problem 05

Evaluate the double integral on the given region

Solution:

$$A=∫42∫42e(8x+3y)dA$$

$$A=\int_2^4 \int_2^4e^{8x}e^{3y}dxdy$$

$$A=\int_2^4e^{8x}dx∗\int_2^4e^{3y}dy$$

$$A=\frac{1}{8}e^{8x}\Big|_3^4∗\frac{1}{3}e^{3y}\Big|_2^4$$

$$A=\frac{1}{24}(e^{32}−e^{16})(e^{12}−e^{6})$$

calculate using R

```{r}
1/24*((exp(32)+exp(16))*(exp(12) - exp(6)))
```