---
title: "FZahir_Assignment10"
author: "Farhana Zahir"
date: "4/05/2020"
output:
  html_document:
    toc: true
    toc_float: yes
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars.A guard agrees to make a series of bets with him. If Smith bets A dollars,he wins A dollars with probability .4 and loses A dollars with probability .6.

Find the probability that he wins 8 dollars before losing all of his money if

# (a) 
He bets 1 dollar each time (timid strategy).

Solution:

$q=.6(Lose)$

$p=.4(Win)$

$q/p=.6/.4=1.5$

$M = 8$

$s=1$

$P(win8)=(1-(q/p)^s)/(1-(q/p)^M=(1-1.5)/(1-1.5^8)$

```{r}
p_win<-(1.5^1-1)/(1.5^8-1)
p_win
```

# (b) 
He bets, each time, as much as possible but not more than necessary tobring his fortune up to 8 dollars (bold strategy).

Solution:

Here Smith bets the entire amount (up to 8). He must win each time or go broke and lose. If he wins, he will follow the following sequence:

1,2,4,8.

He starts with 1 dollar and must win 3 times in a row with probability 0.4. So, the probability of 3 successes in a row is p^4

```{r}
prob<-0.4^3
prob
```

# (c) 
Which strategy gives Smith the better chance of getting out of jail?

Solution

Strategy B (Bold strategy) gives Smith a better chance out of getting out of jail, 0.064> 0.020.
