---
title: "FZahir_Assign14"
author: "Farhana Zahir"
date: "5/08/2020"
output:
  html_document:
    toc: true
    toc_float: yes
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(pracma)
```

This week, we'll work out some Taylor Series expansions of popular functions.

## f(x)=1/1-x


For each function, only consider its valid ranges as indicated in the notes when you are computing the Taylor Series expansion. Please submit your assignment as a R-Markdown document.



\[f\left( x \right) = \frac { 1 }{ (1-x) } \]


\(f(a) + {{ f }^{ \prime }}(a)(x-a) + \frac{{ f }^{ \prime \prime }}{2!}(x-a) + \frac {{ f }^{ \prime \prime \prime }}{3!}(x - a) + \frac {f^{(4)}}{4!}(x - a) +...\)

\(= 1 + 1x + \frac{2}{2!}x^2 + \frac{6}{3!}x^3 + \frac{24}{4!}x^4 +...\)

which reduces to,

$1 + x + x^2 + x^3 + x^4 + ...

```{r}
c = 0
f <- function(x) {1/(1-x)}
a <- taylor(f,x0=c,5)
a
```


Visualizing the approximation function
```{r}
p <- function(x) {a[6] + a[5]*(x-c) + (a[4]*(x-c)^2) + (a[3]*(x-c)^3) + (a[2]*(x-c)^4 +(a[6]*(x-c)^5))}
plot(f, from =-0.25,to=0.75)
par(new=TRUE)
plot(p, from = -0.75,to=0.75,axes=FALSE ,col = 'red')

```

## f(x)=e^x

\[f(x) = e^x\]


\(f(a) + {{ f }^{ \prime }}(a)(x-a) + \frac{{ f }^{ \prime \prime }}{2!}(x-a) + \frac {{ f }^{ \prime \prime \prime }}{3!}(x - a) + \frac {f^{(4)}}{4!}(x - a) +...\)

\(= 1 + x + \frac{x^2}{2} + \frac{x^3}{6} + \frac{x^4}{24} + ...\)


```{r}
c = 0
f <- function(x) {exp(x)}
a <- taylor(f,x0=c,5) 
a
```

Visualizing the approximation function

```{r}
p <- function(x) {a[6] + a[5]*(x-c) + (a[4]*(x-c)^2) + (a[3]*(x-c)^3) + (a[2]*(x-c)^4 +(a[6]*(x-c)^5))}
plot(f, from = -0.5,to=0.75)
par(new=TRUE)
plot(p, from = -0.75,to=0.75,axes=FALSE ,col = 'red')

```

## f(x)=ln(1+x)

\[f(x) = ln(1 + x)\]


\(f(a) + {{ f }^{ \prime }}(a)(x-a) + \frac{{ f }^{ \prime \prime }}{2!}(x-a) + \frac {{ f }^{ \prime \prime \prime }}{3!}(x - a) + \frac {f^{(4)}}{4!}(x - a) +...\)

\(=x - \frac{x^2}{2} + \frac{x^3}{3} - \frac{x^4}{4} +...\)

```{r}
c = 0
f <- function(x) {log(1+x)} 
a <- taylor(f, x0=c, 5)
a
```

Visualizing the approximation function

```{r}
p <- function(x) {a[6] + a[5]*(x-c) + (a[4]*(x-c)^2) + (a[3]*(x-c)^3) + (a[2]*(x-c)^4 +(a[6]*(x-c)^5))}
plot(f, from = -0.25,to=0.75)
par(new=TRUE)
plot(p, from = -0.75,to=0.75,axes=FALSE ,col = 'red')
```