---
title: "FZahir_Assignment09"
author: "Farhana Zahir"
date: "3/29/2020"
output:
  html_document:
    toc: true
    toc_float: yes
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 01 

Chapter 9.3 Exercise 11 page 363

The price of one share of stock in the Pilsdorff Beer Company is given by $Y_{n}$
Finn notices that $X_{n} = Y_{n+1} - Y_{n}$ appear to be independent random variables with $\mu = 0$ and $\sigma  =\frac{1}{2}$
If $Y_{1} = 100$, estimate the probability that:
a) $Y_{365}\geq100$
b) $Y_{365}\geq110$
c) $Y_{365}\geq120$

Solution:

Calculate the expected value and variance for $Y_{365}$

$E(X_{n}) = E(Y_{n+1}-Y_{n}) = E(Y_{n+1})-E(Y_{n})$

$E(X_{i}) = 0 \implies E(Y_{n+1})=E(Y_{n})$

$E(Y_{1}) = E(100) = 100 \implies E(Y_{365}) = 100$

$V(Y_{365}) = V(Y_{364} + X_{365}) = V(Y_{364}) + V(X_{365}) = V(Y_{364}) + \sigma$

$= V(Y_{363}) + 2\sigma = ... = 365\sigma = \frac{365}{4}$

a)
Since $E(Y_{365}) = 100 \implies P(Y_{365} \geq 100) = 0.5$

b)
$\frac{110 - 100}{\sqrt{\frac{365}{4}}} = 1.04685$

```{r}
pnorm(-1.04685)
```

$\implies P(Y_{365}\geq 100) = 0.1476$

c)
$\frac{120 - 100}{\sqrt{\frac{365}{4}}} = 2.0937$

```{r}
pnorm(-2.0937)
```
$\implies P(Y_{365}\geq 100) = 0.01814$


## Problem 02 

Calculate the expected value and variance of the binomial distribution using the moment generating function.

Solution:

For binomial distribution, $P(X=k) = {n \choose k} p^k q^{n-k}$, where $q=1-p$.

The moment generating function is $M_X(t)=(q+pe^t)^n$.

The first moment is $M'_X(t) = n(q+pe^t)^{n-1}pe^t$.

The expected value is the first moment evaluated at $t=0$:

$$
\begin{split}
E(X)=M'_X(0) &= n(q+pe^0)^{n-1}pe^0\\
&= n(q+p)^{n-1}p\\
&= np(1-p+p)^{n-1}\\
&= np1^{n-1}\\
&=np
\end{split}
$$
The second moment is $M''_X(t) = n(n-1)(q+pe^t)^{n-2}p^2 e^{2t}+n(q+pe^t)^{n-1}pe^t$.

Evaluating the second moment at $t=0$:
$$
\begin{split}
E(X^2)=M''_X(0) &= n(n-1)(q+pe^0)^{n-2}p^2 e^0+n(q+pe^0)^{n-1}pe^0\\
&= n(n-1)(1-p+p)^{n-2}p^2+n(1-p+p)^{n-1}p\\
&= n(n-1)p^2+np
\end{split}
$$
The variance is $V(X)=E(X^2)-E(X)^2$:
$$
\begin{split}
V(X) &= n(n-1)p^2+np-n^2p^2 \\
&= np((n-1)p+1-np) \\
&= np(np-p+1-np) \\
&= np(1-p) \\
&= npq
\end{split}
$$
We get the known definitions for binomial distribution - $E(X)=np$ and $V(X)=npq$.


## Problem 03 
Calculate the expected value and variance of the exponential distribution using the moment generating function.

Solution:

For exponential distribution, $f(x)=\lambda e^{-\lambda x}$.

The moment generating function is $M_X(t)=\frac{\lambda}{\lambda-t}, t<\lambda$.

$M'_X(t) = \frac{\lambda}{(\lambda-t)^2}$ and $M''_X(t) = \frac{2\lambda}{(\lambda-t)^3}$.

https://www.youtube.com/watch?v=ROpBZalIt3Q

Expected value:
$$
\begin{split}
E(X)=M'_X(0) &= \frac{\lambda}{(\lambda-0)^2} \\
&= \frac{\lambda}{\lambda^2}\\
&= \frac{1}{\lambda}
\end{split}
$$
Variance:
$$
\begin{split}
V(X) = E(X^2)-E(X)^2 &= M''_X(0)-M'_X(0)^2 \\
&=\frac{2\lambda}{(\lambda-0)^3} - \frac{1}{\lambda^2}\\
&=\frac{2\lambda}{\lambda^3} - \frac{1}{\lambda^2}\\
&=\frac{2}{\lambda^2} - \frac{1}{\lambda^2}\\
&=\frac{1}{\lambda^2}
\end{split}
$$
We get the known definitions for exponential distribution - $E(X)=1/\lambda$ and $V(X)=1/\lambda^2$.