---
title: "FZahir_Assignment04"
author: "Farhana Zahir"
date: "2/23/2020"
output: 
  html_document:
    toc:   true
    toc_float:  true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem Set 1

Verify using R that SVD and Eigenvalues are related

```{r}
A <- matrix(c(1,2,3, -1, 0, 4), nrow = 2, ncol = 3, byrow = TRUE)
```

Compute the left-singular, singular values, and right-singular vectors of A using the svd command.

The vector d contains the singular values
The matrix u contains the left singular vectors (rows).
The matrix v contains the right singular vectors (columns).

```{r}
A_svd <- svd(A, nu=nrow(A), nv=ncol(A))
singular_values <- as.matrix(diag(A_svd$d, nrow=nrow(A), ncol=ncol(A)))
left_singluar_vectors <- as.matrix(A_svd$u)
right_singluar_vectors <- as.matrix(A_svd$v)
```

Singular values from SVD
```{r}
singular_values
```
The left singular vectors from SVD function

```{r}
left_singluar_vectors
```

The right singular vectors from SVD function

```{r}
right_singluar_vectors
```

a) Examine the two sets of singular vectors and show that they are indeed
eigenvectors of X and Y.

Compare left singular vector by computing AAT and taking its eigenvector:

```{r}
X <- A %*% t(A)
X_eigen <- eigen(X) #AAT
X_left_singluar_vectors <- as.matrix(X_eigen$vectors)
X_eigenvalues <- X_eigen$values 
```

let us compare the two

```{r}
left_singluar_vectors
X_left_singluar_vectors
```

The numbers are the same but there are differences in signs. 

The square root of the eigenvalues of X match the singular values of the SVD.

```{r}
singular_values
diag(sqrt(X_eigenvalues), nrow=nrow(A), ncol=ncol(A))
```

Let us now compare right singular vector by computing ATA and taking its eigenvector

```{r}
Y <- t(A)%*% A #ATA
Y_eigen <- eigen(Y)
Y_right_singluar_vectors <- as.matrix(Y_eigen$vectors)
Y_eigenvalues <- Y_eigen$values
```

The SVD right singular vectors and ATA eigenvectors match but there are sign differences

```{r}
right_singluar_vectors
Y_right_singluar_vectors
```

The eigenvalues of X and Y match for nonzero values

```{r}
round(X_eigenvalues,3)
round(Y_eigenvalues,3)
```

## Problem Set 2

```{r}
#The function to get cofactors
getcofactors <- function(M) {
  cofactors <- M
  for(i in 1:nrow(M)){
    for(j in 1:ncol(M)){
      cofactors[i,j] <- (det(M[-i,-j])*(-1)^(i+j)) 
    }
  }
  return(cofactors) # output of cofactors matrix
}

#The function to get the inverse of a square matrix that is invertible
myinverse <- function(M){
  M_cofactors <- getcofactors(M)
  M_cofactors_transpose <- t(M_cofactors)
  M_det <- det(M)
  return(M_cofactors_transpose/M_det)
}
```

Testing the function
```{r}
mat<- matrix(c(4,3,7,8,6,2,2,1,6,1,2,1,2,3,6,7), nrow=4, ncol=4, byrow=TRUE)
mat
```
```{r}
mat_inverse<-myinverse(mat)
mat_inverse
```

Check if mat*mat_inverse=Identity
```{r}
print(mat%*%mat_inverse)
```
