---
title: "FZahir_Assignment05"
author: "Farhana Zahir"
date: "3/01/2020"
output: 
  html_document:
    toc:   true
    toc_float:  true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(pracma)
library(ggplot2)
```
## Problem Set 1

Choose independently two numbers B and C at random from the interval [0, 1] with uniform density. 

Prove that B and C are proper probability distributions. Note that the point (B,C) is then chosen at random in the unit square.

```{r}
#randomly select 10,000 numbers between 0 and 1
B <- runif(10000, min = 0, max = 1)
head(B)
C <- runif(10000, min = 0, max = 1)
head(C)
head(B+C)

```

All values of B and C are between 0 and 1 and sum of the probabilites of each value in B and C is less than 1.

Let us plot the density functions for B and C. We see that both are uniform density functions with values ranging from 0 to 1

```{r}
B_data<-as.data.frame(B)
C_data<-as.data.frame(C)

ggplot(B_data, aes(x=B)) + 
  geom_density() +
  labs(title = "Density Plot of Random Value B (n = 10,000)", x = "Random Values")


ggplot(C_data, aes(x=C)) + 
  geom_density() +
  labs(title = "Density Plot of Random Value C (n = 10,000)", x = "Random Values")


```

Find the probability that

a) B+C < 1/2

$$P(B+C<1/2)=P(X+Y<1/2)=P(0<X<1/2,0<Y<1/2−x)$$

```{r}

a <- sum((B+C) < .5)/10000
print("The probability that B+C will be less than 1/2 is")
print(a)
```

b) BC < 1/2

$$P(B,C<1/2)$$

```{r}
b <- (sum((B*C) < .5)/10000)
print("The probability that B*C will be less than 1/2 is")
print(b)
```


c) |B-C| < 1/2

$$P(|X−Y|< 1/2 , 0<x+y<1) = P(−1/2<X−Y<1/2 ; 0< X< 1−Y)$$

```{r}
c <- sum(abs((B-C)) < .5)/10000
print("The probability that |B-C| will be less than 1/2 is")
print(c)
```

d) max{B,C} < 1/2

$$P(max(B,C)<1/2)=P(B<=1/2,C<=1/2)$$

```{r}
x <- 0
for(i in 1:10000){
  if(max(c(B[i],C[i])) < 0.5){
    x = x+1
  }
}
d <- x/10000
print("The probability that max{B,C} will be less than 1/2 is")
print(d)
```

min{B,C} < 1/2

$$P(min(B,C)<=1/2)=1−P(B>1/2,C>1/2)$$

```{r}
x <- 0
for(i in 1:10000){
  if(min(c(B[i],C[i])) < 0.5){
    x = x+1
  }
}
e <- x/10000
print("The probability that min{B,C} will be less than 1/2 is")
print(e)

```
