---
title: "FZahir_Assignment02"
author: "Farhana Zahir"
date: "2/09/2020"
output: 
  html_document:
    toc:   true
    toc_float:  true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem Set 1

1) Show that ATA!=AAT in general. (Proof and demonstration)

```{r}
#Let us take a 4X3 matrix mat1
mat1<-matrix(1:12,nrow=4, byrow=T)
mat1
#Calculate the transpose of mat1, this is a 3X4 matrix
mat1_transpose<-t(mat1)
mat1_transpose
#mutiply mat1 with mat1_transpose
AAT<-mat1%*%mat1_transpose
AAT
#This is a 4X4 matrix
ATA<-mat1_transpose%*%mat1
ATA
#this is a 3X3 matrix, so AAT is not equal to ATA
```

2) For a special type of square matrix A, we get ATA==AAT. Under what conditions could this be true?

Let us verify if this is true for a 5X5 identity matrix

```{r}
#Construct the identity matrix
Idmat<-diag(5)
Idmat
#Transpose the matrix
IdmatT<-t(Idmat)
AAT1<-Idmat%*%IdmatT
ATA1<-IdmatT%*%Idmat
AAT1==ATA1
#The two are equal, so the conditions are true in case of identity matrices
```

Let us test if this is true for a scalar multiplication of an indentity matrix

```{r}
Idmat1<-5*Idmat
#Using the same steps as above
Idmat1_T<-t(Idmat)
AAT<-Idmat1%*%Idmat1_T
ATA<-Idmat1_T%*%Idmat1
AAT==ATA
#So this works for scalar multiples of identity matrices also
```

Let us test for any square matrix
```{r}
sqmat<-matrix(1:9, nrow=3, byrow=T)
sqmat_t<-t(sqmat)
AAT<-sqmat%*%sqmat_t
ATA<-sqmat_t%*%sqmat
AAT==ATA
det(AAT)
det(ATA)
#This does not work for just any square matrix
```

What about the case of a symmetric matrix (square matrix where a=atranspose)
```{r}
symmat<-matrix(c(1,7,3,7,4,-5,3,-5,6), nrow=3, byrow=T)
symmat_t<-t(symmat)
symmat==symmat_t
AAT<-symmat%*%symmat_t
ATA<-symmat_t%*%symmat
AAT==ATA
#This works for symmetric matrices
```

Lastly we will test a skew symmetric matrix, a square matrix whose ranspose equals to its negative

```{r}
skewmat<-matrix(c(0,2,-45,-2,0,-4,45,4,0), nrow=3, byrow=T)
skewmat_t<-t(skewmat)
AAT<-skewmat%*%skewmat_t
ATA<-skewmat_t%*%skewmat
AAT==ATA
#This also works for skew symmetric matrices
```

So we can see that AAT=ATA for all orthogonal, symmetric and skew symmetric matrices. For reference, go to https://en.wikipedia.org/wiki/Normal_matrix


## Problem Set 2

Write an R function to factorize a square matrix A into LU 

Creating the function

```{r}
LU<-function(mat){
  Lower<-diag(nrow(mat))#start with an identity matrix for lower
  Upper<-mat #upper is built from original matrix using Gaussian elimination
  
  for (c in 1:(nrow(mat) - 1)) { #Iteration stops on row before last
    for (r in (c+1):nrow(mat)) { #Second loop is 1+first, ends in last row
      if(Upper[r,r] != 0)
      Lower[r,c] <- Upper[r,c]/Upper[c,c]
      Upper[r,] <- Upper[r,] - (Upper[c,] * (Upper[r,c]/Upper[c,c]))
  }
}

  LU<-list("L"=Lower, "U"=Upper)
  return(LU)
}

#Help sought from https://rpubs.com/santoshmanjrekar/418200
```

Testing with a 3X3 matrix

```{r}
A<-matrix(1:9, nrow=3, byrow=T)
LUdecomposed<-LU(A)
#Multiply upper and lower matrix
product<-LUdecomposed$L%*%LUdecomposed$U
product==A
##This equals the original 3X3 matrix A
```

Testing with a 4X4 matrix

```{r}
B<-matrix(c(1, 2, -1, -2, 8, 6, -1, -2, 3,-1, -2, 4,5,2,4,1), nrow=4, byrow=T)
LUdecomposed1<-LU(B)
#Multiply upper and lower matrix
product1<-LUdecomposed1$L%*%LUdecomposed1$U
product1==B
#This equals the original 4X4 matrix B
```


