---
title: "FZahir_Discussion13"
author: "Farhana Zahir"
date: "4/29/2020"
output:
  html_document:
    toc: true
    toc_float: yes
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Problem 9 Ex 4.2 page 187

## Problem

A 24 ft ladder is leaning against a house while the base is pulled away at a constant rate of 1 ft/s. At what rate is the top of the ladder sliding down the side of the house when the base is:

(a) 1 foot from the house?
(b) 10 feet from the house?
(c) 23 feet from the house?
(d) 24 feet from the house?

Solutions:

## 1 foot from the house

The ladder and house makes up a right triangle: $x^2+y^2=24^2$

x= the horizontal distance the ladder is from the wall

y= the vertical distance the ladder is up the wall

Taking the derivative with respect to time gives us:

$2x dx/dt + 2y dy/dt = 0$

$\implies dy/dt = -x/y dx/dt$

```{r}
x <- 1
y <- sqrt(24^2-x^2)
dxdt <- 1
dydt <- -x*dxdt/y
dydt
```

The required rate is dy/dt = -0.0417 ft/sec

## 10 feet from the house

Using the derivation in (a) and plugging in the variables

```{r}
x <- 10
y <- sqrt(24^2-x^2)
dxdt <- 1
dydt <- -x*dxdt/y
dydt
```

The required rate is dy/dt = -.458 ft/sec

## 23 feet from the house

Using the derivation in (a) and plugging in the variables

```{r}
x <- 23
y <- sqrt(24^2-x^2)
dxdt <- 1
dydt <- -x*dxdt/y
dydt
```

The required rate is dy/dt = -3.35 ft/sec

## 24 feet from the house

The ladder is completely horizontal and there is no vertical height.

```{r}
x <- 24
y <- sqrt(24^2-x^2)
dxdt <- 1
dydt <- -x*dxdt/y
dydt
```

As the vertical height approaches zero, dy/dt approaches infinity.