---
title: "Discussion9"
author: "Farhana Zahir"
date: "3/25/2020"
output: 
   html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Continuous Independent Trials

Prob 11 Page 363

The price of one share of stock in the Pilsdorff Beer Company is given by $Y_{n}$
Finn notices that $X_{n} = Y_{n+1} - Y_{n}$ appear to be independent random variables with $\mu = 0$ and $\sigma  =\frac{1}{2}$

If $Y_{1} = 100$, estimate the probability that:

a) $Y_{365}\geq100$

b) $Y_{365}\geq110$

c) $Y_{365}\geq120$

Solutions:

First of all, we calculate the expected value and variance for $Y_{365}$

$E(X_{n}) = E(Y_{n+1}-Y_{n}) = E(Y_{n+1})-E(Y_{n})$

$E(X_{i}) = 0 \implies E(Y_{n+1})=E(Y_{n})$

$E(Y_{1}) = E(100) = 100 \implies E(Y_{365}) = 100$

$V(Y_{365}) = V(Y_{364} + X_{365}) = V(Y_{364}) + V(X_{365}) = V(Y_{364}) + \sigma $

$= V(Y_{363}) + 2\sigma = ... = 365\sigma = \frac{365}{4}$

a)

Since $E(Y_{365}) = 100 \implies P(Y_{365} \geq 100) = 0.5$

b)

$\frac{110 - 100}{\sqrt{\frac{365}{4}}} = 1.04685$

```{r}
pnorm(-1.04685)
```

$\implies P(Y_{365}\geq 100) = 0.1476$


c)

$\frac{120 - 100}{\sqrt{\frac{365}{4}}} = 2.0937$

```{r}
pnorm(-2.0937)
```

$\implies P(Y_{365}\geq 100) = 0.01814$