---
title: "605 Discussion01"
author: "Farhana Zahir"
date: "1/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Matrix Multiplication Ex C26, page 191, A first Course in Linear Algebra, Beezer

### Question: Compute the product AB of the two matrices below using Matrix Multiplication(MM)

### Constructing the matrix
```{r}
A<-matrix(c(1,3,1,0,1,0,1,1,2),nrow=3,byrow=TRUE)
A
B<-matrix(c(2,-5,-1,0,1,0,-1,2,1),nrow=3,byrow=TRUE)
B
```

### MM(Matrix Multiplication)

```{r}
C<-A%*%B
C
```

As we can see this results in an identity matrix

We know that A*AInverse will lead to an identity matrix
Let us test if B is inverse of A

```{r}
print(solve(A))
```

This is exactly the matrix B

## Matrix multiplication without using R

Let us look at the step by step multiplication process
We will name variables according to row and column numbers (for example C11=row1column1 of matrixC)

```{r}
C11<-A[1,1]*B[1,1]+A[1,2]*B[2,1]+A[1,3]*B[3,1]
C21<-A[2,1]*B[1,1]+A[2,2]*B[2,1]+A[2,3]*B[3,1]
C31<-A[3,1]*B[1,1]+A[3,2]*B[2,1]+A[3,3]*B[3,1]

C12<-A[1,1]*B[1,2]+A[1,2]*B[2,2]+A[1,3]*B[3,2]
C22<-A[2,1]*B[1,2]+A[2,2]*B[2,2]+A[2,3]*B[3,2]
C32<-A[3,1]*B[1,2]+A[3,2]*B[2,2]+A[3,3]*B[3,2]
  
C13<-A[1,1]*B[1,3]+A[1,2]*B[2,3]+A[1,3]*B[3,3]
C23<-A[2,1]*B[1,3]+A[2,2]*B[2,3]+A[2,3]*B[3,3]
C33<-A[3,1]*B[1,3]+A[3,2]*B[2,3]+A[3,3]*B[3,3]

M<-matrix(c(C11,C21,C31,C12,C22,C32,C13,C23,C33), nrow=3, byrow=TRUE)
M
```

This gives thesame result as using R.