---
title: "Discussion3"
author: "Farhana Zahir"
date: "2/12/2020"
output: 
   html_document:
    toc:   true
    toc_float:  true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Problem C10 page 388

Find the charactertistic polynomial of the matrix 

```{r load-packages, warning = FALSE}
library(knitr)
library(pracma)
```

```{r matrix}
A = matrix(c (1,2,3,4), nrow=2, byrow=T)
print(A)
```

## Solution using R

```{r}
charpoly(A)
```

This gives us the characteristic polynomial as below:

$${ PA(x)=λ}^{ 2 }-5{ λ } -2$$

## Solution without using charpoly

We will find the characteristic polynomial of a 2X2 matrix

$$A=\begin{vmatrix} 1 & 2 \\ 3 & 4\end{vmatrix}$$

We must solve the following to get eigenvalues

$$det(\begin{vmatrix} λ & 0 \\ 0 & λ \end{vmatrix}-\begin{vmatrix} 1 & 2\\ 3 & 4\end{vmatrix})\ =0$$

Subtracting gives us

$$det(\begin{vmatrix} λ-1 & -2 \\ -3 & λ-4\end{vmatrix})= 0$$

Next step is to solve for the determinant using det(A)=ad-bc

$$(λ-1)(λ-4)- (-3)(-2)=0$$

$$λ(λ-4)-1(λ-4)-6=0$$
$$λ^2-4λ-λ+4-6=0$$
$$λ^2-5λ-2=0$$

This is the same equation as the one given by the coordinated from charpoly function.



