---
title: "FZahir_Discussion14"
author: "Farhana Zahir"
date: "5/06/2020"
output:
  html_document:
    toc: true
    toc_float: yes
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Problem 10 Ex 8.8 page 496

## Problem

find a formula for the nth term of the Taylor series of f(x), centered at c, by finding the coefficients of the first few powers of x and looking for a pattern

$$f(x)=ln(1+x)$$


## Solution by hand:

The taylor series expansion is given by the following:

$$
f(x)=\sum _{ n=0 }^{ \infty  }{ \frac { { f }^{ n }(a) }{ n! } (x-a)^{n} } \\
=f(a)+f'(a)(x-a)+\frac{f''(a)}{2!}(x-a)^{2}+\frac{f'''(a)}{2!}(x-a)^{3}+...
$$

Since c=0, we will be evaluating the Maclaurin series

We want to try and find the nth derivative by finding a few derivatives first and observing a pattern that may emerge

$$
f(x)=ln(1+x)\\
f'(x)=(x+1)^{-1}=\frac{1}{(x+1)}\\
f''(x)=-(x+1)^{-2}=-\frac{1}{(x+1)^{2}}\\
f'''(x)=2(x+1)^{-3}=\frac{2}{(x+1)^{3}}\\
f''''(x)=(2)(-3)(x+1)^{-4}=-\frac{(2)(3)}{(x+1)^{4}}\\
f'''''(x)=(2)(-3)(-4)(x+1)^{-5}=\frac{(2)(3)(4)}{(x+1)^{5}}\\
...\\
...\\
...\\
f^{n}(x)=(-1)^{n}(n-1)!(x+1)^{-n}=(-1)^{n}\frac{(n-1)!}{(x+1)^{n}}
$$

Value of derivatives when x=0:

$$
f(0)=ln(1+x)=ln(1+0)=0\\
f'(0)=\frac{1}{(0+1)}=1\\
f''(0)=-\frac{1}{(0+1)^{2}}=-1\\
f'''(0)=\frac{2}{(0+1)^{3}}=2\\
f''''(0)=-\frac{(2)(3)}{(0+1)^{4}}=-(2)(3)\\
f'''''(0)=\frac{(2)(3)(4)}{(x+1)^{5}}=(2)(3)(4)\\
...\\
...\\
...\\
f^{n}(0)=(-1)^{n}\frac{(n-1)!}{(0+1)^{n}}=(-1)^{n}(n-1)!
$$

Computing our nth Taylor Series

$$
T(x)=f(x)=\sum _{ n=0 }^{ \infty  }{ \frac { { f }^{ n }(a) }{ n! } (x-a)^{n} }\\
=\sum _{ n=0 }^{ \infty  }{ \frac { (-1)^{n}(n-1)! }{ n! } (x)^{n} }\\
=\sum _{ n=0 }^{ \infty  }{ \frac { (-1)^{n}(n-1)! }{ n(n-1)! } (x)^{n} }\\
=\sum _{ n=0 }^{ \infty  } (-1)^{n} \frac{x^{n}}{n}
$$

## Chacking approximation using pracma

```{r message=FALSE, warning=FALSE}
library(pracma)
```

```{r}
f <- function(x) log(1+x)
p <- taylor(f, 0, 4)

x <- seq(-1.0, 1.0, length.out=100)
yf <- f(x)
yp <- polyval(p, x)
plot(x, yf, type = "l", main =' Taylor Series Approximation of 1/(1-x)',col = "blue", lwd = 3)
lines(x, yp, col = "red")
legend('topleft', inset=.05, legend= c("Approximation using Taylor Series", "f(x)=1/(1+x)")
       , lwd=c(2.5,2.5), col=c('red', 'blue'), bty='n', cex=.75)
```

Source: https://www.rdocumentation.org/packages/pracma/versions/1.9.9/topics/taylor


## Expansion using rSympy
```{r message=FALSE, warning=FALSE}
library(rSymPy)
```

```{r}
sympy("var('x')")
sympy("var('p')")
sympy("var('1+x')") 
xt <- sympy("p=series(log(1+x), x, 0, 15)")#returns the first 15 orders of the series
print(xt)         
```
