---
title: "FZahir_Assignment07"
author: "Farhana Zahir"
date: "3/15/2020"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=F}
library(tidyverse)
library(ggplot2)
```
## Problem 1

Let X1, X2, . . . , Xn be n mutually independent random variables, each of
which is uniformly distributed on the integers from 1 to k. Let Y denote the
minimum of the Xi’s. Find the distribution of Y.

Solution:

$k^n$ represents the total number of combinations. This will form the denominator to calculate probability.

$(k-1)^n$ represents all combinations where none of the $x_i$ is equal to 1.

$$P(x=1)=\frac{k^n-(k-1)^n}{k^n}$$
$$P(x=2)=$$
If x=3, then there are $k^n-(k-2)^n-[k^n-(k-1)^n]$ different combinations
We can simplify this to
$$k^n-(k-2)^n-[k^n-(k-1)^n]
=k^n-k^n-(k-2)^n+(k-1)^n=(k-1)^n-(k-2)^n$$

$$P(x=2)=\frac{(k-2+1)^n-(k-2)^n}{k^n}$$

Therefore, we can generalize

$$P(x=y)=\frac{(k-y+1)^n-(k-y)^n}{k^n}$$


## Problem 2 part a

Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.)

a)What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a geometric. (Hint: the probability is equivalent to not failing during the first 8 years..)

P(machine will fail) = .10
P(machine will not fail) = .90

$$P(x=n)=p*(1−p)^{n−1}$$
Here, the success factor is the machine will fail, so p=0.10, q=0.90
The probabilty that the machine will fail after 8 years means that the machine will not fail for each of the first 8 years. This is caulated as $q^8$ and is denoted by p below.
I calculated the same using the R function pgeom and find out I have to use an n parameter that is less by 1. This is because as per the source code of pgeom, the formula used is
$$p(x)=p(1-p)^n$$
Therefore, to find the probability that the machine is working by the end of the 8th day, I need to use n parameter of 7.

*https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/Geometric*

*https://rstudio-pubs-static.s3.amazonaws.com/327096_819b7905fc8e4cca83ad7a8bbff24723.html*

```{r}
p<-(0.9^8)
print('the probablity using formula is')
p
prob<-pgeom(7, 0.1, lower.tail = F)
print('The probablity using pgeom is')
prob
```

Let us plot the distribution

```{r}
#still working on 8th day
still_working <-1-pgeom(0:10,.1)
still_working
```

Plotting the Cumulative Density Function

```{r}
n<-7
p<-0.10
data.frame(x = 0:10, 
           cdf = pgeom(q = 0:10, prob = p, lower.tail = F)) %>%
  mutate(Failures = ifelse(x <= n, n, "other")) %>%
ggplot(aes(x = factor(x), y = cdf, fill = Failures)) +
  geom_col() +
  geom_text(
    aes(label = round(cdf,2), y = cdf + 0.01),
    position = position_dodge(0.9),
    size = 3,
    vjust = 0
  ) +
  labs(title = "Cumulative Probability of X = 8 Failures.",
       subtitle = "Geometric Distribution",
       x = "Failures prior to first success (x)",
       y = "probability") 
```

Expected value 

In probability theory and statistics, the geometric distribution is either of two discrete probability distributions:

The probability distribution of the number X of Bernoulli trials needed to get one success, supported on the set { 1, 2, 3, ... }
The probability distribution of the number Y = X − 1 of failures before the first success, supported on the set { 0, 1, 2, 3, ... }

the following form of the geometric distribution is used for modeling the number of failures until the first success:

$$P(Y=k)=(1-p)^{k}p$$
for k = 0, 1, 2, 3, ....
*https://en.wikipedia.org/wiki/Geometric_distribution*

$$E(X)=q/p=\frac{.90}{.10}=9$$

$$StDev(X)=\sqrt\frac{1-p}{p^2}=\sqrt\frac{0.9}{0.1^2}=9.49$$


## Problem 2 part b



What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as an exponential.

Solution:

$$E(x)=\frac{1}{\lambda}$$

$$\sigma=\sqrt\frac{1}{\lambda^2}$$
```{r}
lambda<-0.10
exp_value<-1/lambda
st_dev<-sqrt(1/lambda^2)
exp_value
st_dev

```

Probability that machine will fail after eight years

```{r}
#no of trials
lambda<-0.10
n<-8
p_exp<-pexp(8, lambda, lower.tail = F)
p_exp
#using formula
e<-exp(1)
p1_exp<-e^(-lambda*n)
p1_exp
```

Plotting the distribution

```{r}
n<-8
lambda<-0.10
data.frame(x = 0:10, 
           cdf = pexp(q = 0:10, rate = lambda, lower.tail = F)) %>%
  mutate(Failures = ifelse(x <= n, n, "other")) %>%
ggplot(aes(x = factor(x), y = cdf, fill = Failures)) +
  geom_col() +
  geom_text(
    aes(label = round(cdf,2), y = cdf + 0.01),
    position = position_dodge(0.9),
    size = 3,
    vjust = 0
  ) +
  labs(title = "Cumulative Probability of X = 8 Failures.",
       subtitle = "Exponential Distribution",
       x = "Failures prior to first success (x)",
       y = "probability") 
```


## Problem 2 part c

What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years).

Solution:

$$E(x)=np$$

$$\sigma=\sqrt(npq)$$
```{r}
exp_binom<-n*p
stdev_binom<-sqrt(n*p*(1-p))
exp_binom
stdev_binom
```

Prob machine fails after 8yrs=prob machine does not fail for the first 8 yrs

```{r}
p_binom<-pbinom(0, 8, 0.1)
p_binom
#Using formula P(x<=n)=p^0*q^n
p1_binom<-(p^0)*((1-p)^8)
p1_binom
```

## Problem 2 part d

What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a Poisson.

Solution:

$$E(x)=\lambda, \sigma=\sqrt\lambda$$
$$\lambda=\frac{np}{t}=\frac{8*0.1}{1}=0.8$$
```{r}
lambda<-0.8
exp_pois<-lambda
stdev_pois<-sqrt(lambda)
exp_pois
stdev_pois
```

```{r}

p_pois<-ppois(0,lambda)
p_pois
#Checking with the formula
p1_pois<-exp(-0.8)
p1_pois
```

